Anthropomorphism in Human-Robot Interaction: Nature, Dynamics and Underlying Cognitive Mechanisms

### 2014-11-25 paper.tex
New large pass on the sections up to 3.2.1 + merge Julia's conclusion
Section 5 (measuring anthro.) is now within the conclusion

\newcommand{\AntE}[1][]{%
      \ifthenelse{\isempty{#1}}%
      {$\mathcal{A}^{e}$}
      {$\mathcal{A}^{e}(#1)$}
}

\newcommand{\AntNorm}[1][]{%
      \ifthenelse{\isempty{#1}}%
      {$\widehat{\mathcal{A}}$}
      {$\widehat{\mathcal{A}}(#1)$}
}

@@

\newcommand{\antE}[1][]{%
      \ifthenelse{\isempty{#1}}%
      {\mathcal{A}^{e}}
      {\mathcal{A}^{e}(#1)}
}

\newcommand{\antENorm}[1][]{%
      \ifthenelse{\isempty{#1}}%
      {\widehat{\mathcal{A}^{e}}}
      {\widehat{\mathcal{A}^{e}}(#1)}
}

\newcommand{\ICA}{{$\mathcal{A}_0$~}}
\newcommand{\ICAe}{{$\mathcal{A}^{e}_0$~}}
\newcommand{\SLA}{{$\mathcal{A}_\infty$~}}
\newcommand{\SLAe}{{$\mathcal{A}^{e}_\infty$~}}
\newcommand{\sla}{{\mathcal{A}_\infty}}
\newcommand{\AntMax}{{$\mathcal{A}_{max}$~}}
\newcommand{\AntEMax}{{$\mathcal{A}^{e}_{max}$~}}
\newcommand{\antEMax}{{\mathcal{A}^{e}_{max}}}

@@

\def\journal{Cognitive Science}
\def\DOI{}
\def\articleType{Research Article}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Fink {et~al.}} 
\def\Authors{Julia Fink\,$^{1*\dagger}$, S\'{e}verin Lemaignan$^{1\dagger}$, Claire Braboszcz$^{2}$ and Pierre Dillenbourg$^{1}$}
\def\Address{$^{1}$Computer-Human Interaction in Learning and Instruction (CHILI) \\ Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
\vspace{0.25cm}
$^{2}$Laboratory for Neurology and Imaging of Cognition (LabNIC) \\
Universit\'{e} de Gen\`{e}ve \\ CH-1211 Gen\`{e}ve, Switzerland

@@

\title[Dynamics of Anthropomorphism]{Dynamics of Anthropomorphism in Human-Robot Interaction}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\extraAuth{}
\topic{The Uncanny Valley Hypothesis and Beyond}

@@

By building upon experimental results from several studies, a comprehensive
synthesis of literature, as well as insights from social psychology,
cognitive science, and neuroscience, this article proposes to extend our current
understanding of anthropomorphism. We propose a \textbf{qualitative model} that defines a
clear terminology and accounts for the \textbf{non-monotonic dynamics} of
the phenomenon. We further relate the observed effects of anthropomorphism to
their \textbf{underlying cognitive precursors}.

We examine several \textbf{factors} that determine anthropomorphism, namely the
design of the non-human agent, the individual psychological determinants of the
human user, and the context of use. Moreover, we discuss several episodes of
interaction that influence the anthropomorphic projections on the robot, like
the \textbf{novelty effect}, the role of \textbf{familiarization} and the
impact of \textbf{disruptive behaviors}.

@@

A phenomenon that is often studied related to this is \textbf{anthropomorphism}.
This article aims at fostering a better understanding of
anthropomorphism (noted \Ant in this paper) in human-robot interaction by
considering the phenomenon as a whole. Robotics researchers often tend to
believe that anthropomorphism only describes a static set of human-like features
of a robot (like shape, speech capabilities, facial expressions, etc.). We refer
to these characteristics as the \emph{anthropomorphic design} of the
robot~\citep{fink_anthropomorphism_2012}, and propose that
\emph{anthropomorphism} in general refers to a \emph{social phenomenon} that
emerges from the (real or imagined) \emph{interaction} between a non-human agent
(a robot in our case) and a human~\citep{persson_anthropomorphism_2000}. According
to~\cite{epley_when_2008}, anthropomorphism includes that the human perceives,
for instance, emotional states, motivations, or intentions in the non-human
agent, and tends to ascribe those qualities to it.  As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with artificial agents such as robots.

@@

Figures~\ref{fig:roomba-activities} and~\ref{fig:roomba-perception} give a first
perspective on how naive users interact and perceive robots over a long period
of time. Here, \emph{Roomba} vacuum cleaning robots were given to nine
households for a 6-months period. The aim was to explore how people use and
perceive the robot over time, and how they generally live with the Roomba in
their home. At various time points, users were asked to rate their perception of
different features of the robot on a 7-point rating scale. As
Figure~\ref{fig:roomba-perception} shows, some features remained stable over
time (\emph{ease of use} and \emph{fun}), and others decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one does not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on one's life (and the level of activity pictured in
Figure~\ref{fig:roomba-activities} qualitatively confirms this intuition). Even
if one can argue that the Roomba is not designed to foster interaction, it
raises a first question: how to sustain engagement between a robot and a human
over a long period of time? And, as a pre-requisite to answer, do we actually
understand the psychological phenomenon, and can we effectively account for the
dynamics of these interactions?

@@

Figure~\ref{fig:qualitative-score} calls for other initial remarks. This diagram
shows data from a (yet unpublished) study in which 13 pairs of children (4-5
years old) played for about 20 minutes a domino game with a robot toy box (the
robot was tasked to carry domino pieces from one child to the other,
Figure~\ref{fig:ranger-expe}).  The $x$-axis represents the number of social
interactions between the children and the robot (like talking to the robot or
showing objects), based on video annotations. On the $y$-axis, we have a
synthetic qualitative measurement of anthropomorphic perception of the the
robot, based on two interviews with the children. The scatter plot evidences one
cluster of high-anthropomorphizers/low-interactors and a another group of
children who interact more with the robot and anthropomorphize it less: the
more the children engaged into interaction, the less prone to an anthropomorphic
perception they were.

These results may seem counter-intuitive at first sight, and illustrate that the
mechanisms that relate anthropomorphism to adoption and engagement are
non-trivial. In section~\ref{sec:cognition-neuroscience}, we sketch a model of
the cognitive correlates of anthropomorphism that partially account for these
results.

@@

First, by reviewing what anthropomorphism actually means: besides an
\emph{artifact-centered} understanding that explains anthropomorphism from the
human-like design of the non-human artifact, it seems important to explicit the
social component, namely a \emph{human-centered} understanding, of
anthropomorphism.

@@

While we discuss in section~\ref{meaning} the different meanings of
anthropomorphism found in literature, it seems appropriate to already precisely
define the terminology used in this article.

We use the term \textbf{anthropomorphism} to describe the \emph{psychological}
phenomenon which arises in the \emph{social context} of an interaction between a
human agent and a non-human agent, in our case a robot (see Figure
\ref{fig:anthropofig}). We use this term to refer to both people's tendency to
\textit{perceive} (\textbf{anthropomorphic perceptions}) human-like
characteristics, and to the tendency to \textit{ascribe} or \textit{attribute}
such human-like characteristics to the agent or artifact
(\textbf{anthropomorphic projections}). Besides, we call {\bf anthropomorphic
effects} the \emph{observable manifestations} of anthropomorphism, \ie the
human's behaviors and actions toward a non-human agent that reflect the
underlying cognitive process of anthropomorphizing.

Despite the fact that ``anthropomorphism'' literally means ``human
form'' we propose that the term should \textit{not} be used to refer to the
\textit{form / design} of a non-human agent or artifact. To refer to an
imitation of human-like \textit{form / design} of artificial agents, we propose
to use the term \textbf{anthropomorphic design}. This explicit terminology
prevents the commonly encountered confusion between anthropomorphism as a
psychological, subjective phenomenon, and the external, objective factors that
\emph{influence} the phenomenon itself.

@@

The remainder of this article is organized as follows.  According to the three
perspectives that we take, section \ref{sec:anthropomorphism} establishes a
common ground in the understanding of anthropomorphism. We review literature
from several relevant domains and integrate the different explanations into an
understanding of anthropomorphism as a multi-factor phenomenon.  Based on this,
section~\ref{sec:our-ideas} proposes a new qualitative model of the
\textit{dynamics of anthropomorphism}. This model describes the variances and
dynamics of people's tendency to anthropomorphize an artificial agent, such as a
robot. Section \ref{sec:cognition-neuroscience} then investigates the cognitive
and neuro-biological mechanisms of anthropomorphism to consolidate our
considerations and the proposed model.  Section \ref{sec:measuring} looks at how
these models can be operationalized, and in particular, what are the tools at
hand to measure anthropomorphism. Finally, section \ref{sec:conclusion}
concludes this article and offers starting points for further research on
anthropomorphism and the effect of the human-like design approach in robotics.

@@

Literature on anthropomorphism is quite diverse. There exist different and
partly conflicting definitions of the term anthropomorphism, and it still
appears controversial that humans tend to \textit{``see human in the non-
human''} \citep{epley_seeing_2007}. As anthropomorphism arises in an interaction
between a human and a non-human agent~\citep{persson_anthropomorphism_2000}, one
can also take two main \textit{perspectives} when trying to explain why humans
anthropomorphize~\citep{lee_human_2005}: an artifact-centered perspective and a
human-centered perspective.

After discussing and clarifying the different meanings of anthropomorphism, we
present the main aspects of an \emph{artifact-centered} perspective on
anthropomorphism. We then provide a synthesis of literature that study
 anthropomorphism from the perspective of the human: we aim here at providing a broad
picture of this phenomenon by summarizing theories from developmental
psychology, cognitive science, neuroscience, and social psychology.

We then briefly discuss the impact of the interaction context on
anthropomorphism, before restating anthropomorphism as a \emph{multi-factor
phenomenon}.

\subsection{Overview of the diverse understandings of anthropomorphism}
\label{meaning}

Naturally, conflicts in how terms are used lead to a diversity of understandings
of the phenomenon which is described: anthropomorphism is a phenomenon that is
studied in very different domains that each integrate their own understanding,
use the same term to refer to different things.  Despite the fact that there is
no commonly accepted definition, the terms \textit{anthropomorphism},
\textit{anthropomorphic} or \textit{human-like} are often used as if their
meanings were clear and agreed upon \citep{persson_anthropomorphism_2000}. It is
however argued that these terms might be misused
\citep{duffy_anthropomorphism_2002,epley_when_2008}.  For instance, some
researchers refer to \textit{``the robot's level of anthropomorphism''}
\citep{bartneck_is_2007,feil-seifer_human-robot_2008}, whereas others disagree
to such a statement because they understand that a system or artifact itself
does not \textit{``contain anthropomorphism''} (anthropomorphism is not
intrinsic to the robot) but only gives rise to the process of anthropomorphizing
in a given user and situation. Consequently,
\cite{persson_anthropomorphism_2000} conclude that anthropomorphism emerges in
the \textit{interaction} between the technology and the user, and has to be
understood as such. We come back on this point in the next section.

Another paradox is that, while human's tendency to anthropomorphize is commonly
observed, the phenomenon is at the same time still rather poorly understood
\citep{epley_seeing_2007}. The phenomenon itself has been found to be very
complex, sometimes subtle and as such hard to study
\citep{duffy_anthropomorphism_2002,epley_when_2008}. Besides, it is not clear
how to operationalize and measure anthropomorphism (we discuss this specific question
in section \ref{sec:measuring}).

Depending on which definition of anthropomorphism is applied, not everything
colloquially labeled as \emph{anthropomorphism} really is anthropomorphism.  For
example, it is not always clear whether a person (may it be an adult or a child)
who gives an anthropomorphic response, believes that the other agent really
thinks or acts like a human or whether he/she is using anthropomorphism in a
metaphoric sense \citep{leeds_childrens_1992}.  This call for explicitely
stating what we mean when talking about anthropomorphism.

Several attempts have been made to clarify what the psychological phenomenon
\emph{anthropomorphism} actually is.  From a psychology point of view,
\citet{epley_when_2008} explain what anthropomorphism is \textit{not}. For
instance, in their understanding, anthropomorphism does not include behavioral
descriptions of observable actions but it requires going beyond what is directly
observable. They also suggest that anthropomorphism does not merely entail
\textit{animism}, as animate life is not a \textit{uniquely} human-like
characteristic.\footnote{For the whole list and more details, the reader may
refer to the original work \citep{epley_when_2008}.}

As human-likeness itself is manifold and complex, some researchers argue that
there might not just be one single type of anthropomorphism but different levels
or gradations. For instance, \citet{persson_anthropomorphism_2000} proposes to
understand anthropomorphism as a multi-layered phenomenon, namely
\emph{primitive categorization}, \emph{primitive psychology},
\emph{folk-psychology}, \emph{traits}, \emph{social roles}, and \emph{emotional
anthropomorphism}. Correspondingly, \citet{ruijten_introducing_2014} propose to
understand anthropomorphism as a continuum ranging from low to high human
likeness, and to measure it as such.

@@

\subsection{Artifact-centered perspective on anthropomorphism}

The \emph{artifact-centered} perspective explains anthropomorphism from the
ascertainable parts of the non-human agent, namely the anthropomorphic form of
it. This section deals with the question how far anthropomorphism can be
explained from the design of the artifact.  One of the most prominent working
hypothesis in this field is the Uncanny Valley hypothesis
\citep{mori_uncanny_1970}. In general, it is proposed that human-like designs of
robots can evoke both positive and negative feelings in a human user.

The direction of causality in the artifact-centered view goes from the artifact
to the humans: the artifact (or a certain feature of it) \emph{encourages} the
human to anthropomorphize it.  This perspective assumes that humans directly
(mindlessly) respond to life-like and social cues that the non-human agent or
system emits \citep{nass_machines_2000}. Without thoughtful mental processing,
humans tend to simply apply stereotypes and heuristics to the non-human agent,
and in turn apply human-human social schemas and norms to the occurring
interactions.  \cite{takayama_perspectives_2012} proposes to refer to this
immediate (sometimes visceral) sense in a situation as \textit{in-the-moment},
compared to a more distanced cogitation and consideration which she calls
\textit{reflective}. As this distinction implies different cognitive processes,
it is more approriate to apply a human-centered perspective and we will come
back to it later, in section \ref{sec:cognitive-expl}.

\paragraph{The Physical shape} of a robot refers to `how the robot looks like'
and includes its general form, and other visible features, such as its size, the
materials used, its color, and specific details. Generally, robots can be
divided into human-like (anthropomorphic) and mechanical (non-anthropomorphic)
robots. However, it makes sense to be slightly more specific about the form, and
\cite{fong_survey_2003} suggest four categories: anthropomorphic, zoomorphic,
caricatured, functional. The borders and transitions between these categories
are not very well defined.  Of particular importance seems to be the presence or
absence of specific human body parts, such as arms and hands, legs, a torso or a
head and face. Of those, the faces of robots have received remarkable attention
concerning how they are perceived and accepted. \cite{disalvo_all_2002} studied
what features and dimensions of a humanoid robot's face most dramatically
contribute to people's perception of its humanness. The authors analyzed 48
humanoid robot heads and found that the presence of certain features, the
dimensions of the head, and the total number of facial features heavily
influence the perception of humanness in robot heads.

\paragraph{Expression, communication, and behavior} refers to how the robot
expresses its human qualities through interacting and behaving. In fact, many
humanoid robots express their human-likeness not in physical shape but only
through the \textit{suggestion} of features~\citep{disalvo_all_2002}. Besides
this, communication and interaction modalities of a robot are likely to play a
role to what extent the robot will be anthropomorphized. Can the user interact
with the robot through a touch interface, or using speech? Does the robot
respond using linguistic utterances or non-linguistic sounds? Is the voice
human-like or synthetic, on-board or coming from an external second device? A
study by~\cite{takayama_im_2009} found that this has an effect on people's
perception of trust in the robot and how convincing the robot appears to them.
Besides this, we expect that a robot's behavioral aspects concerning
human-social interaction have an impact. Does the robot use polite formulations,
etiquette, appropriate facial expressions? Are its responses and reactions
timely? Usually, robots that directly interact with humans are slow, which
appears not very human-like to most participants.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		HUMAN-CENTERED PERSPECTIVE: SOCIAL PHENOMENON
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Human-centered perspective on anthropomorphism}

In contrast to the artifact-centered perspective, in the human-centered
perspective the argumentation is that the human is self-motivated to
anthropomorphize the non-human artifact. As already mentioned, both these views
go hand in hand.

Humans tend to perceive human-like characteristics such as physical appearance,
emotional states, or inner mental states and motivations in non-human agents
that do not have these characteristics themselves, such as animals, natural
forces, religious agents, technological gadgets, or mechanical devices
\citep{epley_when_2008}.\footnote{According to \citet{epley_when_2008}, a
non-human agent can be anything that acts -- or is believed to act -- with
apparent independence.} One common reason for anthropomorphism is that by
perceiving some human-like characteristics in the non-human agent, this agent is
made more \emph{graspable}, \emph{understandable}, \emph{predictable}, and one
can feel \emph{more empathetic} toward it, and possibly one can then anticipate
the agent's behavior. These are central motivations when trying to explain
anthropomorphism, and we will come back to them in more detail later on.
\citet{eddy_attribution_1993} explain that, because of these reasons, people
commonly anthropomorphize their pets, by ascribing mental states, intentions or
feelings to them.\footnote{Interestingly, pet-ownership seems to have a
significant impact on human-robot interaction: several studies suggest that
pet-owners are more likely to anthropomorphize technologies and robots than
non pet-owners} But how can these motivations behind anthropomorphism be
explained from a human-centered point of view?

In the following, we present what we found in reviewing literature that applies
a human-centered perspective on anthropomorphism by building on: \textit{(1)} a
line of research in developmental psychology that studies the attribution of
animacy; \textit{(2)} a line of research in more cognitive sciences that studies
the cognitive processes that underlie anthropomorphism, and \textit{(3)} a
theory that explains anthropomorphism based on psychological factors.

@@

The phenomenon of attributing animacy and intentions to nonliving objects such
as simple shapes has been intensively studied in developmental psychology. As
mentioned earlier, Piaget found that children tend to ascribe life to the
nonliving. They also tend to interpret physical phenomena in terms of intention
on the part of the non-human subject. For instance children are likely to argue
that \textit{``the sun is hot because it wants to make people warm''}
\citep{leeds_childrens_1992}. In this example it is in fact questionable whether
children really anthropomorphize the sun (believing that it has an intention) or
whether they are using anthropomorphism in a metaphoric sense (because their
conceptions of the world and of the living are not yet fully formed). However,
experiments by \cite{heider_experimental_1944} and
\cite{michotte_perception_1963} showed that also adult participants attribute
animacy and intention to simple things and shapes based on motion. In both
experiments, participants viewed animations of simple shapes, such as circles or
triangles. Asked to describe what they observed, most people developed elaborate
stories, attributing motivations, emotions and relationships between the
objects.\footnote{For more details, the reader may refer to the ``attribution
theory''.}

>>>>

\newcommand{\AntE}[1][]{%
      \ifthenelse{\isempty{#1}}%
      {$\widetilde{\mathcal{A}}$}
      {$\widetilde{\mathcal{A}}(#1)$}
}

@@

\newcommand{\antE}[1][]{%
      \ifthenelse{\isempty{#1}}%
      {\widetilde{\mathcal{A}}}
          {\widetilde{\mathcal{A}}(#1)}
}

\newcommand{\antENorm}[1][]{%
      \ifthenelse{\isempty{#1}}%
      {\widehat{\mathcal{A}}}
      {\widehat{\mathcal{A}}(#1)}
}

\newcommand{\IPA}{{$\mathcal{A}_0$~}}
\newcommand{\IPAe}{{$\widetilde{\mathcal{A}_0}$~}}
\newcommand{\SLA}{{$\mathcal{A}_\infty$~}}
\newcommand{\SLAe}{{$\widetilde{\mathcal{A}_\infty}$~}}
\newcommand{\sla}{{\mathcal{A}_\infty}}
\newcommand{\AntMax}{{$\mathcal{A}_{max}$~}}
\newcommand{\AntEMax}{{$\widetilde{\mathcal{A}_{max}}$~}}
\newcommand{\antEMax}{{\widetilde{\mathcal{A}_{max}}}}

@@

\def\journal{Cognitive Science}
\def\DOI{}
\def\articleType{Research Article}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Fink {et~al.}} 
\def\Authors{Julia Fink\,$^{1*\dagger}$, S\'{e}verin Lemaignan$^{1\dagger}$, Claire Braboszcz$^{2}$ and Pierre Dillenbourg$^{1}$}
\def\Address{$^{1}$Computer-Human Interaction in Learning and Instruction
    (CHILI) \\ \'{E}cole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
\vspace{0.25cm}
$^{2}$Laboratory for Neurology and Imaging of Cognition (LabNIC) \\
Universit\'{e} de Gen\`{e}ve \\ CH-1211 Gen\`{e}ve, Switzerland

@@

\title[Dynamics of Anthropomorphism]{Anthropomorphism in Human-Robot
Interaction: Nature, Dynamics and Underlying Cognitive Mechanisms}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\extraAuth{}
\topic{The Uncanny Valley Hypothesis and Beyond}

@@

By building upon experimental results from several studies, a comprehensive
synthesis of literature, as well as insights from social psychology, cognitive
science, and neuroscience, this article proposes to extend our current
understanding of anthropomorphism. We examine the several \textbf{factors} that
determine anthropomorphism, namely the design of the non-human agent, the
individual psychological determinants of the human user, and the context of use.
Moreover, we discuss several episodes of interaction that influence the
anthropomorphic projections on the robot, like the \textbf{novelty effect}, the
role of \textbf{familiarization} and the impact of \textbf{unexpected
behaviors}.

We then propose a \textbf{qualitative model} that defines a clear terminology
and accounts for the \textbf{non-monotonic dynamics} of the phenomenon. We
further relate the observed effects of anthropomorphism to their
\textbf{underlying cognitive precursors}.

@@

A phenomenon that is often studied related to this is \textbf{anthropomorphism}.
This article aims at fostering a better understanding of anthropomorphism (noted
\Ant in this paper) in human-robot interaction by considering the phenomenon as
a whole. Robotics researchers often tend to consider that anthropomorphism only
describes a static set of human-like features of a robot (like its shape, its
speech capabilities, facial expressions, etc.). We refer to these
characteristics as the \emph{anthropomorphic design} of the
robot~\citep{fink_anthropomorphism_2012}, and propose that
\emph{anthropomorphism} in general refers to a \emph{social phenomenon} that
emerges from the (real or imagined) interaction between a non-human agent (a
robot in our case) and a human~\citep{persson_anthropomorphism_2000}.  According
to~\cite{epley_when_2008}, anthropomorphism includes the human perception of
emotional states, motivations, intentions in the non-human agent, and tends to
ascribe those qualities to it.  As such, the dynamic and socio-cognitive
dimensions of anthropomorphism are essential to understand the complex bonds
that human users build with artificial agents such as robots. This article aims
at advancing our understanding of this phenomenon, both in term of an
interdisciplinary literature synthesis, and in term of qualitative yet formal
models that will support further research in this field.

@@

Figures~\ref{fig:roomba-activities} and~\ref{fig:roomba-perception} give a first
perspective on how naive users interact and perceive robots over a long period
of time. \emph{Roomba} vacuum cleaning robots were given to nine
households for a 6-months period. The aim was to explore how people use and
perceive the robot over time, and how they generally live with the Roomba in
their home. At various time points, users were asked to rate their perception of
different features of the robot on a 7-point rating scale. As
Figure~\ref{fig:roomba-perception} shows, some features remained stable over
time (\emph{ease of use} and \emph{fun}), and others decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one does not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on one's life (and the level of activity pictured in
Figure~\ref{fig:roomba-activities} qualitatively confirms this intuition). Even
if one can argue that the Roomba is not designed to foster interaction, it
raises a first question: how to sustain engagement between a robot and a human
over a long period of time? And, as a pre-requisite to answer, do we actually
understand the psychological phenomenon, and can we effectively account for the
dynamics of these interactions?

@@

Figure~\ref{fig:qualitative-score} calls for other initial remarks. This diagram
shows data from a (yet unpublished) study in which 13 pairs of children (4-5
years old) played for about 20 minutes a domino game with a robot toy box. The
robot was tasked to carry domino pieces from one child to the other
(Figure~\ref{fig:ranger-expe}).  The $x$-axis represents the number of social
interactions between the children and the robot (like talking to the robot or
showing objects), based on video annotations. On the $y$-axis, we have a
synthetic qualitative measurement of anthropomorphic perception of the the
robot, based on two interviews with the children. The scatter plot evidences one
cluster of high-anthropomorphizers/low-interactors and a another group of
children who interact more with the robot and anthropomorphize it less: the more
the children anthropomophized, the less engaged into the interaction they were.

These results may seem counter-intuitive at first sight, and illustrate that the
mechanisms that relate anthropomorphism to adoption and engagement are
non-trivial. In section~\ref{sec:cognition-neuroscience}, we will sketch a model of
the cognitive correlates of anthropomorphism that partially account for these
results.

@@

First, by reviewing what anthropomorphism actually means: besides an
\emph{artifact-centric} understanding that explains anthropomorphism from the
human-like design of the non-human artifact, it seems important to explicit the
social component, namely a \emph{human-centric} understanding, of
anthropomorphism.

@@

Despite the fact that there is no commonly accepted definition, the terms
\textit{anthropomorphism}, \textit{anthropomorphic} or \textit{human-like} are
often used as if their meanings were clear and agreed
upon~\citep{persson_anthropomorphism_2000}. It is however argued that these
terms might be misused~\citep{duffy_anthropomorphism_2002,epley_when_2008}, and
while some researchers would refer for instance to \textit{``the robot's level
of anthropomorphism''} \citep{bartneck_is_2007,feil-seifer_human-robot_2008},
others disagree and argue that a system or artifact itself does not
\emph{contain anthropomorphism} (anthropomorphism is not intrinsic to the robot)
but only \emph{gives rise} to the process of anthropomorphizing in a given user
and situation.

In the frame of this article, we refer to the term \textbf{anthropomorphism} to
describe the \emph{psychological} phenomenon which arises in the \emph{social
context} of an interaction between a human agent and a non-human agent, in our
case a robot. We use this term to refer to both people's tendency to
\textit{perceive} (\textbf{anthropomorphic perceptions}) human-like
characteristics, and to the tendency to \textit{ascribe} or \textit{attribute}
such human-like characteristics to the agent or artifact
(\textbf{anthropomorphic projections}). Besides, we call {\bf anthropomorphic
effects} the \emph{observable manifestations} of anthropomorphism, \ie the
human's behaviors and actions toward a non-human agent that reflect the
underlying cognitive process of anthropomorphizing.

Despite the fact that ``anthropomorphism'' literally means ``human form'' we
propose that the term should \textit{not} be used to refer to the \emph{form}
and/or \emph{design} of a non-human agent or artifact. To refer to an imitation
by an artificial agent of an human-like form, we propose to use the term
\textbf{anthropomorphic design}. This explicit terminology prevents the commonly
encountered confusion between anthropomorphism as a psychological, subjective
phenomenon, and the external, objective factors that \emph{influence} the
phenomenon itself.

@@

The remainder of this article is organized as follow.  According to the three
perspectives that we take, section \ref{sec:anthropomorphism} establishes a
common ground in the understanding of anthropomorphism. We review literature
from several relevant domains and integrate the different explanations into an
understanding of anthropomorphism as a multi-factor phenomenon.  Based on this,
section~\ref{sec:our-ideas} proposes a new qualitative model of the dynamics of
anthropomorphism. This model describes the variances and dynamics of people's
tendency to anthropomorphize an artificial agent, such as a robot. Section
\ref{sec:cognition-neuroscience} then investigates the cognitive and
neuro-biological mechanisms of anthropomorphism to consolidate our
considerations and the proposed model.  Section \ref{sec:measuring} looks at how
these models can be operationalized, and in particular, what are the tools at
hand to measure anthropomorphism. Finally, section \ref{sec:conclusion}
concludes this article and offers starting points for further research on
anthropomorphism and the effect of the human-like design approach in robotics.

@@

Anthropomorphism arises in an interaction between a human and a non-human agent,
and one can consequently take two main perspectives when trying to explain why
humans anthropomorphize~\citep{lee_human_2005}: an artifact-centric perspective
and a human-centric perspective. We present these two perspectives, with a
special focus on the complex human aspects: we attempt there to summarize the
main approaches from developmental psychology, cognitive science, neuroscience,
and social psychology. We then briefly discuss the impact of the interaction
context on anthropomorphism, which will finally lead us to globally restate
anthropomorphism as a \emph{multi-factor phenomenon}.

@@

\subsection{Artifact-centric perspective on anthropomorphism}

The \emph{artifact-centric} perspective explains anthropomorphism from the
ascertainable parts of the non-human agent, namely the anthropomorphic form of
it. One of the most prominent working hypothesis in this field is the Uncanny
Valley hypothesis~\citep{mori_uncanny_1970}. It suggests that human-like designs
of robots can evoke both positive and negative feelings in a human user.

The direction of causality in the artifact-centric view goes from the artifact
to the humans: the artifact (or a certain feature of it) \emph{encourages} the
human to anthropomorphize it.  This perspective assumes that humans directly
(mindlessly) respond to life-like and social cues that the non-human agent or
system emits \citep{nass_machines_2000}. Without thoughtful mental processing,
humans tend to simply apply stereotypes and heuristics to the non-human agent,
and in turn apply human-human social schemas and norms to the occurring
interactions.  \cite{takayama_perspectives_2012} proposes to refer to this
immediate (sometimes visceral) sense in a situation as \textit{in-the-moment},
compared to a more distanced cogitation and consideration which she calls
\textit{reflective}. As this distinction implies different cognitive processes,
it is more approriate to apply a human-centric perspective and we will come
back to it later, in section \ref{sec:cognitive-expl}.

\paragraph{The Physical shape} of a robot refers to ``how the robot looks like''
and includes its general form, and other visible features, such as its size, the
materials used, its color, and specific details. Specifically,
\cite{fong_survey_2003} suggest four categories: anthropomorphic, zoomorphic,
caricatured, functional. The borders and transitions between these categories
are not very well defined. The presence or absence of specific human body parts,
such as arms and hands, legs, a torso or a head and face seems of particular
importance. Of those, the faces of robots have received remarkable attention:
\cite{disalvo_all_2002} studied for instance what features and dimensions of a humanoid
robot's face most dramatically contribute to people's perception of its
humanness. The authors analyzed 48 humanoid robot heads and found that the
presence of certain features, the dimensions of the head, and the total number
of facial features heavily influence the perception of humanness in robot heads.

\paragraph{Expression, communication, and behavior} refers to how the robot
expresses its human qualities through interacting and behaving, and as a matter
of fact, many humanoid robots primarily express their human-likeness through the
\textit{suggestion} of features~\citep{disalvo_all_2002} instead of their
physical shape. Interaction modalities also play a role to what extent the robot
will be anthropomorphized. Can the user interact with the robot through a touch
interface, or using speech? Does the robot respond using linguistic or
non-linguistic utterances? Is the voice human-like or synthetic, on-board or
coming from an external second device? A study by~\cite{takayama_im_2009} found
that this has an effect on people's perception of trust in the robot and how
convincing the robot appears to them.  Finally, the implementation of \emph{social
rules} into the robot's behavior also impacts the tendency to anthropomorphize. Does the
robot use polite formulations, etiquette, appropriate facial expressions? Are
its responses and reactions timely? As it stands today, robots tend on the
contrary to interact with humans at a slow pace, which does not elicit a
human-like perception by the user.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		HUMAN-centric PERSPECTIVE: SOCIAL PHENOMENON
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Human-centric perspective on anthropomorphism}

Humans tend to perceive human-like characteristics such as physical appearance,
emotional states, or inner mental states and motivations in non-human agents
that do not have these characteristics themselves, such as animals, natural
forces, religious agents, technological gadgets, or mechanical devices
\citep{epley_when_2008}.\footnote{According to \citet{epley_when_2008}, a
non-human agent can be anything that acts -- or is believed to act -- with
apparent independence.} One common explanation for anthropomorphism is that by
perceiving some human-like characteristics in the non-human agent, this agent is
made more \emph{graspable}, \emph{understandable}, \emph{predictable}, and one
can feel \emph{more empathetic} toward it, and possibly one can then anticipate
the agent's behavior. These are the typical reasons used for instance
by~\citet{eddy_attribution_1993} to explain people's tendency to
anthropomorphize their pets, by ascribing mental states, intentions or feelings
to them.\footnote{Interestingly, pet-ownership seems to have a significant
impact on human-robot interaction: several studies suggest that pet-owners
are more likely to anthropomorphize technologies and robots than non
pet-owners}

In the following, we present what we found in reviewing literature that applies
a human-centric perspective on anthropomorphism by building on: \textit{(1)} a
line of research in developmental psychology that studies the attribution of
animacy; \textit{(2)} a line of research in more cognitive sciences that studies
the cognitive processes that underlie anthropomorphism, and \textit{(3)} a
theory that explains anthropomorphism based on psychological factors.

@@

The phenomenon of attributing animacy and intentions to nonliving objects such
as simple shapes has been intensively studied in developmental psychology.
Piaget found that children tend to ascribe life to the nonliving and they also
tend to interpret physical phenomena in terms of intention on the part of the
non-human subject. For instance children are likely to argue that \textit{``the
sun is hot because it wants to make people warm''} \citep{leeds_childrens_1992}.
It is questionable whether children really anthropomorphize the sun (believing
that it has an intention) or whether they are using anthropomorphism in a
metaphoric sense (because their conceptions of the world and of the living are
not yet fully formed). However, experiments by \cite{heider_experimental_1944}
and \cite{michotte_perception_1963} showed that also adult participants
attribute animacy and intention to simple things and shapes based on motion. In
both experiments, participants viewed animations of simple shapes, such as
circles or triangles. Asked to describe what they observed, most people
developed elaborate stories, attributing motivations, emotions and relationships
between the objects.\footnote{For more details, the reader may refer to the
\emph{attribution theory}.} People tend to attribute animacy and intention
to moving shapes, even when there is no similarity of the shape to a human. This
tendency to animize / anthropomorphize is already developed (and even more
distinct) in infants and children. Thus besides the anthropomorphic cues emitted
by the design of the artifact, also developmental stages and people's natural
reaction play a role how far we interpret movements (as one of the
most basic ``life-like'' cues) for instance.
